---
title: "Corruption and human development"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: journal
    highlight: zenburn
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(ggplot2)
```


# Initial play with the data
Loading the data 
```{r}
housing <- read.csv(here('data', 'landdata-states.csv'))
head(housing)
```

Creating a histogram

```{r}
hist(housing$Home.Value)
```
```{r}
ggplot(housing, aes(x = Home.Value)) +
  geom_histogram()
```

 Base colored scatter plot
```{r}
plot(Home.Value ~ Date,
     data=subset(housing, State == "MA"))
points(Home.Value ~ Date, col="purple",
       data=subset(housing, State == "TX"))
legend(1975, 400000,
       c("MA", "TX"), title="State",
       col=c("black", "purple"),
       pch=c(1, 1))
```
 
 Using ggplot
 
```{r}
ggplot(subset(housing, State %in% c("MA", "TX")),
       aes(x=Date,
           y=Home.Value,
           color=State))+
  geom_point()
```

## Scatter plots

```{r}
hp2001Q1 <- subset(housing, Date == 2001.25) 
ggplot(hp2001Q1, aes(y = Structure.Cost, x = log(Land.Value))) +
  geom_point()
```

Adding prediction line


```{r}
hp2001Q1$pred.SC <- predict(lm(Structure.Cost ~ log(Land.Value), data = hp2001Q1))

plt <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

plt + geom_point(aes(color = Home.Value)) +
  geom_line(aes(y = pred.SC))
```

Adding labels to the points

```{r}
plt <- ggplot(hp2001Q1, aes(x = log(Land.Value), y = Structure.Cost))

plt + geom_text(aes(label=State), size = 3)
```

## Aesthetic Mapping VS Assignment
Note that variables are mapped to aesthetics with the aes() function, while fixed aesthetics are set outside the aes() call. 

## Mapping variables to other aesthetics
```{r}
plt + geom_point(aes(color = Home.Value, shape = region))
```

### Statistical transformations

```{r}
housing.sum <- aggregate(housing["Home.Value"], housing["State"], FUN = mean)

ggplot(housing, aes(x = State, y = Home.Value)) +
  geom_bar(stat = 'identity')
```

# Scales

## Scales: Controlling aesthetics mapping

Aesthetic mapping (i.e., with `aes()`) only says that a variable should be mapped to an aesthetic. It doesn’t say how that should happen. For example, when mapping a variable to shape with `aes(shape = x)` you don’t say what shapes should be used. Similarly, `aes(color = z)` doesn’t say what colors should be used. Describing what colors/shapes/sizes etc. to use is done by modifying the corresponding scale. In ggplot2 scales include

- position
- color and fill
- size
- shape
- line type
Scales are modified with a series of functions using a scale_<aesthetic>_<type> naming scheme. Try typing scale_<tab> to see a list of scale modification functions.

# Common Scale Arguments
The following arguments are common to most scales in ggplot2:
 
- name: the first argument gives the axis or legend title
- limits: the minimum and maximum of the scale
- breaks: the points along the scale where labels should appear
- labels: the labels that appear at each break

Specific scale functions may have additional arguments; for example, the `scale_color_continuous` function has arguments `low` and `high` for setting the colors at the low and high end of the scale.

## Scale Modification Examples
Start by constructing a dotplot showing the distribution of home values by Date and State.


```{r}
p1 <- ggplot(housing, 
             aes(x = State, y = Home.Price.Index)) +
  theme(legend.position = "top",
        axis.text = element_text(size = 6))
(p2 <- p1 + geom_point(aes(color = Date),
                       alpha = 0.6, size = 1.5,
                       position = position_jitter(width = 0.25, height = 0)))
```


Change the breaks for the  x axis and color scales

```{r}
p2 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"))
```

Changing the colour schemes

```{r}
p2 + scale_x_discrete(name="State Abbreviation") +
  scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = "blue", high = "purple")
```

```{r}
library(scales)
p2 + scale_color_continuous(name="",
                         breaks = c(1976, 1994, 2013),
                         labels = c("'76", "'94", "'13"),
                         low = muted("blue"), high = muted("red"))
```

## Using different color scales

`scale_color_gradient2` will interpolate between three different colors

```{r}
p2 + scale_color_gradient2(name="",
                        breaks = c(1976, 1994, 2013),
                        labels = c("'76", "'94", "'13"),
                        low = muted("blue"),
                        high = muted("red"),
                        mid = "gray60",
                        midpoint = 1994)
```

# Faceting

- Faceting is ggplot2 parlance for small multiples
- The idea is to create separate graphs for subsets of data
- ggplot2 offers two functions for creating small multiples:
- `facet_wrap()`: define subsets as the levels of a single grouping variable
- `facet_grid()`: define subsets as the crossing of two grouping variables
Facilitates comparison among plots, not just of geoms within a plot

```{r}
p5<- ggplot(housing, aes(x = Date, y =Home.Value))

p5 + geom_line(aes(color = State))
```

## Faceting

```{r}
(p5 <- p5 +geom_line(aes(color = State)) +
  facet_wrap(~State, ncol = 10) +
   theme(legend.position ="None"))
```


## Themes

```{r}
p5 + theme_dark() +
  theme( text =  element_text(color = "turquoise"))
```

# Creating and saving new themes

```{r}
theme_plex <- function(base_size = 9,
                       strip_text_size = 12,
                       strip_text_margin = 5,
                       subtitle_size = 13,
                       subtitle_margin = 10,
                       plot_title_size = 16,
                       plot_title_margin = 10,
                       ...) {
    ret <- ggplot2::theme_minimal(base_family = "IBMPlexSans",
                                  base_size = base_size, ...)
    ret$strip.text <- ggplot2::element_text(hjust = 0, size=strip_text_size,
                                            margin=margin(b=strip_text_margin),
                                            family="IBMPlexSans")
    ret$plot.subtitle <- ggplot2::element_text(hjust = 0, size=subtitle_size,
                                               margin=margin(b=subtitle_margin),
                                               family="IBMPlexSans")
    ret$plot.title <- ggplot2::element_text(hjust = 0, size = plot_title_size,
                                            margin=margin(b=plot_title_margin),
                                            family="IBMPlexSans-Bold")
    ret$axis.text.y = element_text(colour = "purple")
    ret
}

p5 + theme_plex() + theme(legend.position = "None")
```



