---
title: "Musical visualization"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: journal
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Loading the libraries

```{r, message=FALSE}
library(tidyverse)
library(here)
library(tuneR)
```

# Getting the data
I found that [Kunst der Furge](http://www.kunstderfuge.com/bach/chamber.htm#Cello) has a number of MIDIS for classical music including some of J.S Bach (my all times favourite). So I did not hesitate to jump and analyse this data when I realized it was available.


```{r}
if (!file.exists(here('data','bach.mid')))
  download.file('http://www.kunstderfuge.com/-/midi.asp?file=bach/suite_for_unaccompanied_cello_bwv-1007_1_(c)grossman.mid', here('data', 'bach.mid'))
```

MIDI files contain each note to be played for a song and should contain all information I need for the viz. 

The `tuneR` package allows you to read a MIDI file and get a nice data frame in return. The package even gives individual `track`s with names, meaning that you can figure out which instrument plays each note.

So the first thig to do is to load the whole MIDI in a single data frame.

```{r}
songs <-readMidi('./data/bach.mid')

head(songs)
```

This can be easily splitted into individual tracks and notes

```{r}
tracks <- songs %>%
  filter(event == 'Sequence/Track Name') %>%
  transmute(track, track_name = parameterMetaSystem)

notes <- getMidiNotes(songs) %>%
  inner_join(tracks, by = 'track') %>%
  arrange(time) %>%
  head(50)
```

